---
title: "Reshaping"
author: "Becky Tang"
date: "9/26/2022"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "math118-slides.css"
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(magick)
knitr::opts_chunk$set(warning = F,message = F,
                      fig.align = "center") 
data("iris")
```

class: center, middle

## Housekeeping

---

class: center, middle

# Reshaping

---

# Why reshape?

- Sometimes, our data is not always in the format that we would like

- We want to preserve all the information, but "massage" it prior to analysis

  - Move information between rows and columns
  


---

# Types of data frames

- Data frames are often described as **wide** or **long**

- **Wide**: when a row has more than one observation, and the units of observation (e.g., individuals, countries, households) are on one row each

- **Long**: when a row has only one observation, but the units of observation are repeated down a column

---

# Long to wide data

- This is our original data, where `country`  is our unit of observation.

- Why is this considered long data?


```{r echo = F}
set.seed(1)
country_long <- data.frame(
    expand.grid(country = c("Sweden", "Denmark", "Norway"), month  = c("jan","feb", "march")),
    avgtemp = round(runif(9, 3, 12), 0)
    )

country_long
```


---

# Long to wide data

- Want to reshape from to wide data, where each unit of observation (`country`) has exactly one row


.pull-left[
Long data:
```{r echo = F}
country_long
```
]

.pull-right[
Wide data:
```{r echo = F}
country_long %>%
  pivot_wider(names_from = month, values_from = avgtemp) 
```

]



---

# Long to wide data

- `pivot_wider()` function turns data from long to wide

- Takes two arguments:

  - `names_from`: the name of the variable(s) in the data frame to get the name of the output column
  
  - `values_from`: the name of the variable(s) in the data frame to get the cell values from

---

# Long to Wide data

.pull-left[
Original long data:
```{r echo = F, eval = T}
country_long 
```
]

.pull-right[
Desired wide data:
```{r echo = F}
country_long %>%
  pivot_wider(names_from = month, values_from = avgtemp) 
```

```{r eval = F}
country_long %>%
  pivot_wider(names_from = ____,
              values_from = _____)
```

.question[What should I pass in for `names_from` and `values_from`?]
]


---
# Long to Wide data

```{r eval = T}
country_long %>%
  pivot_wider(names_from = month, #<<
              values_from = avgtemp) #<<
```

---

# Wide to long data

- Long data structure is often preferred

- Often required for advanced statistical analysis and graphing.
  

.pull-left[
Now this is our original data: 

```{r echo = F}
country_wide <- data.frame(
    country = c("Sweden", "Denmark", "Norway"),
    "jan" = country_long$avgtemp[1:3],
    "feb" = country_long$avgtemp[4:6],
    "march" = country_long$avgtemp[7:9])
country_wide 
```
]

.pull-right[
And this is what we want:

```{r echo = F}
country_wide %>%
  pivot_longer(cols = 2:4, names_to = "month", values_to = "avg_temp")
```
]

---

# Wide to Long data

- `pivot_longer()` function can be used to go from wide to long data

- Takes three arguments:

  - `cols`: specify which columns in the dataframe to pivot into longer format. That is, the ones that should "move"
  
  - `names_to`: a string (your choice) specifying the name of the new column created from the column names species by `cols`
  
  - `values_to`: a string (your choice) specifying the name of the column to create from the data stored in cell values



---

# Wide to Long data

This is my original data. 

```{r}
country_wide
```

For each row, I want a country, the month, and the temperature within that month. What should I choose for the arguments `cols`?

--

- `cols`: the columns `jan`, `feb`, and `march`

--

- `names_to`: anything you'd like, so long as it makes sense. I will choose "month"

- `values_to`: anything you'd like, so long as it makes sense. I will choose "avg_temp"

---

# Wide to Long data

```{r}
country_wide %>%
  pivot_longer(cols = c(jan, feb, march), #<<
               names_to = "month", #<<
               values_to = "avg_temp") #<<
```

---

# Wide to Long data

- Can specify column indices in `cols` argument

- What you choose for `values_to` or `names_to` doesn't affect resulting structure

.pull-left[
```{r}
country_wide %>%
  pivot_longer(cols = 2:4, #<< 
               names_to = "month",
               values_to = "temp") #<<
```
]

.pull-right[
```{r}
country_wide %>%
  pivot_longer(cols = -country, #<< 
               names_to = "month",
               values_to = "temp")
```

]



