---
title: "Strings"
author: "Becky Tang"
date: "10/7/22"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "math118-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 3, fig.width = 5, dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center") 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```


```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r echo = FALSE}
states <- c("alabama", "alaska", "arizona", "arkansas", "california", 
            "colorado", "connecticut", "delaware", "florida", "georgia", 
            "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", 
            "kentucky", "louisiana", "maine", "maryland", "massachusetts", 
            "michigan", "minnesota", "mississippi", "missouri", "montana", 
            "nebraska", "nevada", "new hampshire", "new jersey", 
            "new mexico", "new york", "north carolina", "north dakota", "ohio", 
            "oklahoma", "oregon", "pennsylvania", "rhode island",
            "south carolina", "south dakota", "tennessee", "texas", "utah", 
            "vermont", "virginia", "washington", "west virginia", "wisconsin",
            "wyoming")
```


## Housekeeping

- Homework 4 due Tuesday, Oct. 11 at 11:59pm

- This is an application exercise today (the last one before the midterm!)

- You are welcome to come to any of my office hours next week:

  - Monday: 10:30am-12:00pm
  - Tuesday: 11:00am-12:00pm
  - Wednesday: 10:30am-12:00pm

- Please fill out this anonymous survey: https://forms.gle/jcdJhoFRUHcx4vsR9


---

class: middle, center

## `stringr`

---

## `stringr`

In addition to the `tidyverse`, we will use the package `stringr`.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(stringr)
```

`stringr` provides tools to work with character strings.

- Functions in `stringr` have consistent and memorable names

- All begin with `str_` (`str_count`, `str_detect`, `str_trim`, etc)

- All take a vector of strings as their first argument


---


## Preliminaries

Character strings in R are defined by double quotation marks. 

They can include letters, numbers, punctuation, whitespace, etc.

```{r}
string1 <- "MATH 118 is my favorite class."
string1
```

You can combine character strings in a **vector**.

```{r}
string2 <- c("MATH 118", "Data Science", "Middlebury")
string2
```

---

## To quote or not to quote?

```{r eval = F}
datascience %>%
  filter(Major == "Computer Science")
```

Why does `Major` not have quotes, but `"Computer Science"` does?

--

- If you do not use quotes, R assumes you are referring to the name of an object

- If you use quotes, R assumes you are simply entering a character string 

---

## To quote or not to quote?

```{r}
#This defines an object 
object_name <- 2

# This is the name of an object
object_name

# This is a character string
"object_name"
```

---

## Include a quotation in a string?

Why doesn't the code below work?

.midi[
```{r eval = FALSE}
string3 <- "I said "Hello" to my class"
```
]

--

To include a double quote in a string, *escape it* using a backslash `\`.

--

.midi[
```{r}
string4 <- "I said \"Hello\" to my class."
```
]

--

What if you want to include an actual backslash?

--

.midi[
```{r}
string5 <- "\\"
```
]

This may seem tedious but it will come up later!


---

## U.S. States

To demonstrate functions from `stringr` we will use a vector of all 50 states.

.midi[
```{r}
states
```
]

---

## `str_length`

Given a string, return the number of characters.

.midi[
```{r}
string1 <- "Math 118 is my favorite class."
str_length(string1)
```
]

Given a vector of strings, return the number of characters in each string.

.midi[
```{r}
str_length(states)
```
]

--

.pull-left[
- Alabama: 7
- Alaska: 6
- Arizona: 7
- Arkansas: 8
]
.pull-right[
- California: 10
- Colorado: 8
- Connecticut: 11
- ...
]
 

---

## `str_c`

Combine two or more strings.

```{r}
str_c("Math 118", "is", "my", "favorite", "class")
```


--

Use `sep` to specify how the strings are separated.

```{r}
str_c("Math 119", "is", "my", "favorite", "class", sep = "-")
```

```{r}
str_c("Math 119", "is", "my", "favorite", "class", sep = " ")
```

---

## `str_c`

Combine together the strings within a character vector using `collapse`:

```{r}
string_vec <- c("Math", "118", "is", "my", "favorite", "class")
str_c(string_vec, collapse = "")
```

```{r}
str_c(string_vec, collapse = " ")
```

--

What happens if you don't use `collapse`?

```{r}
str_c(string_vec)
```

---

## `str_to_lower` and `str_to_upper`

Convert the case of a string from lower to upper or upper to lower.

.midi[
```{r}
str_to_upper(states)
```
]

---

## `str_to_title`

Converts a string into a title-cased version: converts the first character of each words to uppercase and the remaining characters in lowercase

```{r}
str_to_title("math is 188 my favorite class")
```

---

## `str_sub`

Extract subsets (parts) of a string from `start` to `end`, inclusive.

.midi[
```{r}
str_sub(states, 1, 4)
```
]

--

.midi[
```{r}
str_sub(states, -4, -1)
```
]

---

## `str_sub` and `str_to_upper`


Can combine `str_sub` and `str_to_upper` to capitalize each state.

.midi[
```{r eval = F}
str_sub(states, 1, 1) <- str_to_upper(str_sub(states, 1, 1))
```
]

.question[What is this code doing?]

--
```{r echo = F}
str_sub(states, 1, 1) <- str_to_upper(str_sub(states, 1, 1))
```

```{r}
states
```

---

## `str_sort`

.question[What do we think this code is doing?]

.midi[
```{r eval = F}
str_sort(states, decreasing = TRUE)
```
]

--

Sort a vector of strings, in decreasing alphabetical order.

```{r echo = F}
str_sort(states, decreasing = TRUE)
```

---

## Regular Expressions

A .vocab[regular expression] is a sequence of characters that allows you to 
describe string patterns. We use them to search for patterns. 

- extract a phone number from text data
- determine if an email address is valid
- determine if a password has the required number of letters, characters, and symbols
- count the number of times "statistics" occurs in a corpus of text
- ...

---

## Regular Expressions

To demonstrate, we will use a vector of all of the states bordering North 
Carolina.

```{r}
vt_neighbors <- c("Vermont", "New York", "Massachusetts", "New Hampshire")
vt_neighbors
```

---

## Basic Match

We can match exactly. 


```{r}
str_view_all(vt_neighbors, "ew")
```


---

## Basic Match

Match any character using `.`


```{r}
str_view_all(vt_neighbors, "e.")
```

---

## Extract matches

Pulls the match (if it exist) from each element in the character

```{r}
str_extract(vt_neighbors, "New")
```

---

## Anchors

Match the start of a string using `^`

```{r}
str_view_all(vt_neighbors, "^V")
```

---

## Anchors

Match the end of a string using `$`


```{r}
str_view_all(vt_neighbors, "s$")
```

---

## `str_detect`

Determine if a character vector matches a pattern.

```{r}
vt_neighbors
str_detect(vt_neighbors, "a")
```


---

## `str_subset`

Select elements from the character vector that match a pattern.

```{r}
str_subset(vt_neighbors, "e$")
```


---

## `str_count`

How many matches are there in a string?


```{r}
vt_neighbors
str_count(vt_neighbors, "a")
```


---

## `str_replace`

Replace first match with new strings.


```{r}
str_replace(vt_neighbors, "s", "-")
```


---

## `str_replace_all`

Replace all matches with new strings.

```{r}
str_replace_all(vt_neighbors, "s", "-")
```

--

Don't forget to save results:

```{r}
vt_neighbors
vt_neighbors_replaced <- str_replace_all(vt_neighbors, "s", "-")
vt_neighbors_replaced
```


---

## Many Matches

The regular expressions below match more than one character.

- Match any single digit using `\d` or `[[:digit:]]`
- Match all digits using `\d+` or `[[:digit:]]+`
- Match any whitespace using `\s` or `[[:space:]]`
- Match f, g, or h using `[fgh]` 
- Match anything but f, g, or h using `[^fgh]`
- Match lower-case letters using `[a-z]` or `[[:lower:]]`
- Match upper-case letters using `[A-Z]` or `[[:upper:]]`
- Match alphabetic characters using `[A-z]` or `[[:alpha:]]`

Remember these are regular expressions! To match digits you'll need to *escape*
the string, so use `"\\d"`, not `"\d"`

---

## Working within pipeline

```{r echo = F}
vt_df <- data.frame(state = vt_neighbors, state_code = c("VT", "NY", "MA", "NH"))
```

```{r}
vt_df %>%
  mutate(state_code = str_to_lower(state_code))
```

---

## Working within pipeline

```{r}
vt_df %>%
  mutate(state = str_replace_all(state, "[ea]", "-"))
```


---

## Additional resources

- `stringr` website: https://stringr.tidyverse.org/
- `stringr` [cheat sheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/strings.pdf)
- Regular Expressions [Cheat Sheet](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf)
- [Chapter 14: Strings](https://r4ds.had.co.nz/strings.html#matching-patterns-with-regular-expressions) in R for Data Science