---
title: "Spatial data & visualization"
subtitle: "Part 2: Leaflet"
author: "Becky Tang"
date: "12/5/2022"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "math118-slides.css"
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 2, fig.width = 5, 
                     dpi = 200, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center") 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(sf)
library(spData)
```

---

class: middle, center

## Leaflet

---

## Leaflet

- Leaflet is a popular open-source JavaScript library for *interactive* maps

- We can access it via the `leaflet` package:

```{r}
library(leaflet)
```

---

## How to us Leaflet

We can create a Leaflet map with the following steps:

  1. Create a map *widget* by calling `leaflet()`
  
  2. Add layers (i.e. features) by using layer functions (e.g. `addPolygons`)
  
  3. Print the map widget to display it!
  
--

In the following, note that we use `%>%` rather than `+` because we are *not* creating plots with `ggplot`!

---

## Example: Using OpenStreetMaps

```{r}
leaflet() #<<
```

---

## Example: Using OpenStreetMaps

```{r}
leaflet() %>% 
  addTiles() # Add default OpenStreetMap map tiles #<<
```

---

## Example: Using OpenStreetMaps

```{r}
leaflet() %>% 
  addTiles() %>%  # Add default OpenStreetMap map tiles 
  addMarkers(lng=-73.175, lat=44.010, popup="Warner Hall") #<<
```


---

## Airport data

- Recall in last class we created `airports_count` data frame, which held the number of flights that flew to each airport from NYC


```{r echo = F}
library(nycflights13)
data(airports)
data(flights)
dest_count <- flights %>%
  count(dest, name = "count") 
airports_count <- airports %>%  
  inner_join(dest_count, by=c("faa" = "dest"))
rm(flights)
```

```{r}
airports_count %>%
  slice(1:5)
```

---

## Leaflet for airport data

```{r}
leaflet(airports_count) %>% 
  addTiles() %>%  
  addMarkers(lng = ~lon, lat = ~lat, popup = ~faa)  #<<
``` 

---

## Leaflet for airport data

```{r}
leaflet(airports_count) %>% 
  addTiles() %>%  
  addCircleMarkers(lng=~lon, lat=~lat, popup=~faa, 
                   radius = ~count/1000, #<<
                   stroke = FALSE, fillOpacity = 0.5)
```

---

## Leaflet for airport data

```{r}
leaflet(airports_count) %>% 
  addTiles() %>%  
  addCircleMarkers(lng=~lon, lat=~lat, popup=~faa, 
                   radius = ~count/1000, stroke = FALSE, fillOpacity = 0.5) %>%
  setView(-100, 40, zoom = 4) #<<
```

---

## Change base map

```{r}
leaflet(airports_count) %>% 
  addProviderTiles(providers$Stamen.Watercolor) %>%  #<<
  addCircleMarkers(lng=~lon, lat=~lat, popup=~faa, 
                   radius = ~count/1000, stroke = FALSE, fillOpacity = 0.5) %>%
  setView(-100, 40, zoom = 4) 
```


---

class: center, middle

# YOUR TURN!
