---
title: "HW 05 - Bootstrap estimation"
subtitle: "due Tuesday, October 25 at 11:59pm"
output: 
  tufte::tufte_html:
    highlight: pygments
    css: "./math118-hw.css"
    toc: true
    toc_depth: 2
link-citations: yes
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
library(tidyverse)
library(tufte)
library(knitr)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  digits = 2
  )
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction and Data

The goal of this week's homework is to practice creating bootstrap confidence intervals, and visualizing bootstrap distributions. 

# Tips 

- Use a small number of reps (about 100) as you write and test out your code. Once you have finalized all of your code, increase the number of reps to at least 1000 (10,000 if possible) to produce your final results. 
- Write your code and narrative in a reproducible way, so you can easily change the number of reps. Tips on making your code more reproducible:

### Tip 1: Set your seed

When dealing with randomness (as often the case in simulation in statistics), it is important to specify which pseudo-random draw you used in your analysis, so that you or someone else can reproduce the exact numbers you initially report. The `set.seed()` function in R allows you to ensure that all of your analysis relies on a specific pseudo-random draw:

```{r set_seed, eval = F}
set.seed(42)
```

### Tip 2: Define global parameters

Often, we rely on specific parameters values throughout our analysis, and at a later point, we may want to replace them. In order to minimize the need to change your code later, we can assign the parameter values to a name, and use the name (rather than the hard-coded value) downstream. Then, to update your code at a later point, you can just change the value. Here, we are assinging the number of reps to a variable called `num_reps`:

```{r global_params, eval = F}
num_reps <- 100
boot_dist <- diamonds %>%
  infer::specify(response = price) %>%
  infer::generate(reps=num_reps, type = "bootstrap") %>%
  infer::calculate(stat="mean")
```


# The data

```{r include = F}
library(tidyverse)
library(infer)
```

The data may be found by cloning your **hw-05-boston-**  repository available in the [GitHub course organization](https://github.com/math118-fall2022).

Today's data comes from the city of Boston, courtesy of the U.S. Census Bureau. In particular, the `Boston` dataset contains data about median value of owner-occupied housing units in 506 suburbs of Boston. "Owner-occupied housing units" is defined as: one-family houses on less than 10 acres without a business or medical office on the property. The variables and their definitions are as follows:

  - `crim`: per capita crime rate by town 
  - `rm`: average number of rooms per dwelling
  - `age`: proportion of owner-occupied units built prior to 1940
  - `dis`: weighted mean of distances to five Boston employment centers
  - `tax`: full-value property-tax rate per $10,000
  - `ptratio`: pupil-teacher ratio by town
  - `lstat`: lower status of the population (percent)
  - `medv`: median value of owner-occupied homes (in $1000s)

You may load in the data with the following code, where `____` should be
replaced by a meaningful name of your choosing. Don't forget to set `eval = TRUE` before knitting:


```{r eval = FALSE}
___ <- read.csv("data/Boston.csv")
```

# Exercises

Write all R code according to the style guidelines discussed in class. Make sure that your plots have appropriate labels and titles.

```{marginfigure}
**Hint**: Don't forget to set a seed in order to ensure reproducibility!
```

1. Based on this data set, provide a point estimate for the population mean of the median value of owner-occupied homes in Boston, `medv`. 

```{marginfigure}
```

2. Construct a 95%  bootstrap interval for the mean of the median value of owner-occupied homes in Boston. Use at least 1,000 bootstrap samples. Make sure your interval is *reproducible*.

```{marginfigure}
```

3. Visualize the bootstrap distribution and your confidence interval from 
Exercise 2. Interpret the confidence interval you constructed. 

```{marginfigure}
```

4. Considering towns with a pupil-teacher ratio of at most 15, provide a point estimate of the *proportion* of owner-occupied houses in Boston with median values over $40,000.

```{marginfigure}
```

5. For towns with a pupil-teacher ratio of at most 15, construct a **99%**  bootstrap interval for the proportion of owner-occupied
houses in Boston with median values over $40,000.  Make sure your interval is *reproducible*.

```{marginfigure}
```

6. Visualize the bootstrap distribution and your confidence interval from 
Exercise 5. Interpret the confidence interval you constructed.


```{marginfigure}
**Hint**: To simulate the correlation between two variables, use `specify(var1 ~ var2)`. Remember that correlation is still a numerical quantity, so that should help you choose the type of simulation you want to perform.
```

7. Provide a point estimate of the *correlation* between the median housing value of owner-occupied houses in Boston `medv` and the pupil-teacher ratio in the town.

```{marginfigure}
```

8. Construct a 95% bootstrap interval for the correlation between the median housing value of owner-occupied houses in Boston `medv` and the pupil-teacher ratio in the town. Make sure your interval is *reproducible*.

```{marginfigure}
```

9. Construct and report your **99%** bootstrap interval using the bootstrap distribution from Exercise 8. Then answer the following questions:
  
  - How does the 99% bootstrap interval compare to the 95% bootstrap interval calculated in the previous exercise?
  
  - In general, how does the bootstrap interval change when the confidence level increases?
    
```{marginfigure}
```

# Submission

Knit to PDF to create a PDF document. Knit and commit all remaining changes, 
and push your work to GitHub. Make sure all files are updated on your GitHub 
repo. Please only upload your PDF document to Canvas.
