---
title: "RStudio and GitHub Tour"
subtitle: "Tour of RStudio + RMarkdown + GitHub"
date: "9/14/2022"
output: 
  html_document: 
    theme: readable
---

## Getting started

- Go to the course GitHub organization: https://github.com/math118-fall2022. 

- If you see a pending invite at the top of the page, accept the invite to join the course invite. We'll come back to GitHub later.

## Tour of R and RStudio 

Let's take a tour of R and RStudio. I'll do a short demo, then give you a few minutes to try out yourself. 

### Concepts introduced

- Console
- Using R as a calculator
- Environment
- Loading and viewing a data frame
- Accessing a variable in a data frame
- R functions

### Your turn!

- Go ahead and open up RStudio. 

- Use R as a calculator by typing the following in the **console**. 

```{r eval = F}
3 * 5 + 10
```

```{r eval = F}
x <- 3
x + x^2
```

```{r eval = F}
x <- 1:10
x*7
```

## Git and GitHub

### Concepts introduced:
 
- Clone a GitHub repo and make a new RStudio project 
- Configure Git
- Working with a local and remote repository
- Making a change locally, committing, and pushing
- Making a change on GitHub and pulling

### Your turn!

#### Clone a GitHub repo and make a new RStudio project 

- Create a new folder on your Desktop called "Math118"

- Go to the [course organization](https://github.com/math118-fall2022) on GitHub  and find the repo with the prefix **appex-01-**. 

- Clone the repo and make a new project in RStudio by going to `File` in the top left -> `New Project...` -> `Version Control` -> `Git`. Paste in the copied URL in the first blank space "Repsitory URL". Then, under "Create project as a subdirectory of...", hit "Browse" and find your newly created Math118 folder. Then hit "Create Project".

- You should see the files for the Flint Water Crisis analysis in the **Files** pane. 


#### Configure Git

- We will need to allow RStudio to communicate with Github. We will do this by utilizing the `usethis` package. **Packages** are units of reproducible `R` code, and often include reusable `R` functions. Think of packages like phone apps. We only download and open apps that we need. 

  -The equivalent of downloading an app here is *installing* and package. You only need to do this once; installing the same package repeatedly can cause issues. In your **Console**, type:

```{r eval = F}
install.packages("usethis")
```

  - You will see some red text flying across the Console. If you eventually see some black text beginning with "The downloaded binary packages are in...", then you are good!


- Next, we will "open the app", i.e. *load* the package. This is achieved using the `library()` function. In your Console, run the following code:

```{r eval  = F}
library(usethis)
```

  - The Console might hang for a bit, but if you see a new line with a blue `>`, then the code has run successfully!

- Configure git by running the following code in the Console. Fill in your GitHub username and the email address associated with your GitHub account. **Note**: you must keep the quotation marks!

```{r eval  = F}
use_git_config(user.name = "GitHub username", user.email="your email")
```


#### Making a change locally, committing, and pushing

- Open the RMarkdown file **ae01-flint.Rmd**. We will use that for the next part of class. 

- The `R` package that we will rely on most heavily is the `tidyverse` package. Please install it now by typing the following in your Console:

```{r eval = F}
install.packages("tidyverse")
```

- Put your name for the author, then knit the document. *It's important to knit the document, as this saves your progress!*

- Select/check **all** of the updated files in the Git pane, click the "Commit" button, and write the commit message "Updated author name". 

- Click to **push** your changes (you may have to enter your GitHub password).

- If you refresh your repo on GitHub, you will see the updated files with your commit message. 

#### Making a change on GitHub and pulling

- On the corresponding repository in GitHub, find the README.md file at the bottom. Edit this file by clicking on the pencil and typing your name + major + year. 

- Once you're finished editing, hit the green "Commit changes" button.

- Back in RStudio, navigate to the Git tab and **pull** the changes by hitting the down arrow. A small window should pop up with some text.

- Find and open the README.md file in the Files pane. You should see your changes here locally!

## Flint Water Crisis analysis

Follow along with the instructions in **ae01-flint.Rmd**. An html version of the file can be found [here](https://sta199-summer2021.github.io/website/appex/ae01-flint.html).


## Submitting application exercises 

**Once you have completed the activity, push your final changes to your GitHub repo! Make sure your repo is updated on GitHub, and that's all you need to do to submit application exercises for participation.**
