---
title: "AppEx 17: Paris paintings"
date: "11/30/2022"
output: 
  html_document:
    theme: readable
---

## Announcements

Your project Github repositories have been created. Reminder that your project proposal is due next Friday, July 9 at 11:59pm. Use the file `project-proposal.Rmd` to write your proposal.


## Exercises 

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
```

```{r load-data, message = FALSE}
paintings <- read.csv("data/paris_paintings.csv", 
                            na = c("n/a", "", "NA")) %>%
  dplyr::select(Height_in, Width_in, landsALL) %>%
  na.omit()
```

We will be looking at data about Paris Paintings in today's application exercise.
These data were collected by Hilary Coe Cronheim and Sandra van Ginhoven (Duke Art, Art History & Visual Studies PhD students) as part of the Data Expeditions project sponsored by [iiD](https://bigdata.duke.edu/). These data consist of the physical and artistic features of paintings featured in printed catalogues of 28 auction sales in Paris, 1764-1780. In total, these students analyzed 3,393 paintings, their prices, and descriptive details from sales catalogues over 60 variables

We'll primarily focus on the variables: 

- `Height_in`: Height (in inches)
- `Width_in`: Width (in inches)
- `landsALL`: If any type of landscape is mentioned (either lands_sc, lands_figs, or lands_ment)

### Exercise 1

Right now, `landsALL` is coded as a numeric variable in the dataset. Modify the dataset to make `landsALL` a factor so R treats it as categorical variable.


### Exercise 2

Create a scatterplot to visualize the relationship between width and height. Color the points based on the whether the painting has any landscape elements. 



### Exercise 3

Based on your scatterplot, does the relationship between Height and Width differ between paintings with landscape elements and those without? Briefly explain.  

### Exercise 4

Fit a model using width and whether the painting has landscape elements to predict the height. Present the model in a tidy format.



### Exercise 5

Using your model from the previous exercise, 

- Interpret the intercept. 
- Interpret the coefficient (slope) of `Width_in`. 
- Interpret the coefficient (slope) of `landsAll`.

### Exercise 6

Write the equation of the model for 

- paintings with no landscape elements. 
- paintings with landscape elements. 

How do the slopes compare between the two models? How do the intercepts compare? 

### Exercise 7

Now let's see how well the model fits the data. Use the `glance` function to calculate $R^2$ for you model.



### Exercise 8

Interpret the $R^2$.
