---
title: "AE 04: Data Wrangling"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
editor_options: 
  chunk_output_type: console
---


## Clone a repo + start a new project

- Go to the `appex-04-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. 

- Click on the green **Code** button, select the second option "Open with GitHub Desktop"

- The GitHub Desktop application will open up, with a white window that says "Clone a Repository". **Important**: in the second line that says "Local Path", there is a button that says `Choose...`. Click on it, and find and select the folder we created for this course. Then hit the blue `Clone` button.

- After successfully cloning, the window will disappear and you will see the that Current Repository is the one you just cloned. Success!

- Navigate to the project folder you just created within the Math118 folder, and open the `appex04-ncbike.Rmd` file to begin.

## Practice with data wrangling

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

The data set contains all North Carolina bike crash data from 2007-2014. The data was originally downloaded on September 6, 2018 and is available in the dsbox R package. 

Use the code below to load the data from the .csv file

```{r load-data, warning = FALSE, message = FALSE}
ncbikecrash <- read_csv("data/ncbikecrash.csv")
```


### Part 1

**Display the first 10 rows of the ncbikecrash data set.**

```{r}
```


### Part 2 

**Filter for bike crashes in Durham County where the driver age group is 0 to 19.**

```{r}
```

### Part 3

**On which day of the week did the most bike crashes occur? Use the `count` function to get the number of bike crashes for each day (`crash_day`). Display the results in descending order.**

```{r}
```

### Part 4

**Calculate the average estimated crash hour, `crash_hour` (<u>not</u> driver_est_speed) based on whether the crash was a hit and run (`hit_run`).**

```{r}
```