---
title: "Appex 11: Hypothesis Testing"
subtitle: "Part 1"
date: "10/26/2022"
output: 
  html_document:
    theme: readable
---


## Clone a repo + start a new project

Clone the **appex-11** repo on [GitHub](https://github.com/math118-fall2022) and start a new project in RStudio.

We will be using the `asheville` dataset we saw in class last week, which contains data about Airbnb rates in Asheville, NC. You may load in the dataset with the
following code (be sure to set `eval` to be `TRUE` in the following R chunks!):

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(infer)
```


```{r load-data, message = FALSE, echo = FALSE}
asheville <- read_csv("data/asheville.csv")
```

We will begin with Professor Tang demo-ing code. Then we will live-code together, and then you must complete the remainder of the Application Exercise on your own!


## LIVE CODE 

### Exercise 1: discuss with someone next to you

Suppose you are interested in whether the mean price per guest per night is actually less than $80. Choose the correct null and alternative hypotheses.

a. $H_0: \mu = 80 \text{ vs. }H_a: \mu \neq 80$
b. $H_0: \mu = 80 \text{ vs. }H_a: \mu < 80$
c. $H_0: \mu = 80 \text{ vs. }H_a: \mu > 80$
d. $H_0: \bar{x} = 80 \text{ vs. }H_a: \bar{x}  \neq 80$
e. $H_0: \bar{x} = 80 \text{ vs. }H_a: \bar{x}  < 80$
f. $H_0: \bar{x} = 80 \text{ vs. }H_a: \bar{x}  > 80$


### Exercise 2

Let's use simulation-based methods to conduct the hypothesis test specified in Exercise 1. We'll start by generating the null distribution. 

Fill in the code and uncomment the lines below to generate the null distribution. 

```{r set-seed}
set.seed(101)
```


```{r ex-2}
null_dist <- asheville # %>%
  
```

### Exercise 3

Fill in the code and uncomment the lines below to calculate the p-value using the null distribution from Exercise 2.

```{r}
mean_ppg <- asheville # %>% 
  
```

```{r}
null_dist <- asheville # %>%
```


## Exercise 4 

Use the p-value from Exercise 3 to make your conclusion using a significance level of 0.05. Remember, the conclusion has 3 components 

- How the p-value compares to the significance level 
- The decision you make with respect to the hypotheses (reject $H_0$ /fail to reject $H_0$)
- The conclusion in the context of the research problem 

## YOUR TURN!

### Exercise 5

Suppose you are interested in whether the **median** price per guest per night is
equal to or less than $80. Carry out a similar analysis to that
undertaken in Exercises 1 - 4 to test these hypotheses. 


i) State your hypotheses (in words is fine):

```{r null-dist}
# ii) generate null distribution


```

```{r p-val}
# iii) obtain p-value


```

iv) Conclusion: 


