---
title: "AppEx 13: Wine"
date: ""
output: 
  html_document:
    theme: readable
---

# What makes a good wine?

The goal of this application is to continue practicing two-sample statistical inference using
both simulation-based approaches. Use the lecture notes, readings, 
and application exercises to help you complete homework.

# The Data

Today's dataset has been adapted from a large data set of wine reviews from
[Kaggle](https://www.kaggle.com/zynicide/wine-reviews). Reviews were scraped from 
WineEnthusiast, which contain information about both the wine itself as well as 
the reviewer's description of that wine. Here, I provide a subset of the data. 
You may consider each of these observations to be an
independent, representative sample of all wines.

The variables are as follows

- `country`: the country that the wine is from (France, Italy, US)
- `points`: the number of points WineEnthusiast rated the wine on a scale of 1-100. 
- `price`: cost for a bottle (USD)
- `variety`: type of grapes used to make the wine
- `type`: red or white wine
- `title`: title of the wine review
- `description`: reviewer's description of the wine

### Clone a repo + start a new project

- Clone the **appex-13-diff-means** repo and start a new project in RStudio. 


## Load data and packages 

```{r load-packages-data, eval = FALSE}
library(tidyverse)
library(infer)
___ <- read_csv("data/wines.csv")
```

## Exercises

Suppose you are  interested in how the type of wine (red or white) impacts its price. Comprehensively evaluate the hypothesis that
the average price of red wines is different from the average price of white wines in these three countries. Use $\alpha = 0.05$ as the significance level.

### Exercise 1: State your hypotheses

### Exercise 2: Obtain your null distribution

Because we will be simulating randomly, we should first set a seed. Set the seed equal to the exercise number.

```{r seed}
# set seed here

```

```{r ex2}
# obtain null distribution
```

### Exercise 3: Visualize

Display a visualization of your simulated null distribution, and describe the values that would cause you to reject your null hypothesis. Does the observed sample statistic lie in this rejection region?

### Exercise 4: Conclude

What is your p-value, decision, and conclusion in context of the
research question?

### Exercise 5: Confidence interval

Construct and interpret a 95% two-sided confidence interval for the
difference you investigated in Exercise 2.


