---
title: "AE 09: Strings"
name: "Your name"
output: html_document
date: "10/07/2022"
---

## Clone a repo + start a new project

Go to the `appex-09-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. 


```{r setup, warning=FALSE, message = F}
library(tidyverse)
addresses <- c("24 Merchants Row", "30 Seymour St.", "3 Mill St.", "111 Maple St.", "7 Bakery Ln.")
```

### Exercise 1

Create your own string vector called `name_parts` which holds at least two elements: your first name, your middle name (if you have one), and your last name. Please type them all in **lowercase** letters.

*Remember*: a vector is created using the `c()` function, where the components are separated by commas.

```{r}
```

### Exercise 2

Combine the elements of `name_parts` into a single string separated by spaces. Name this new string `my_name`. 

```{r ex2}
```

### Exercise 3

Capitalize the first letter of each character in `name_parts`.

```{r ex3}

```

Now we will work a vector called `addresses`:

```{r addresses}
addresses
```

### Exercise 4

Obtain the number of characters in each string of `addresses`.

```{r ex4}


```

### Exercise 5

Select the elements from `addresses` that end in "St.".

```{r ex5}

```

### Exercise 6: Optional stretch goal!

Pull or access only the numeric part of `addresses` (i.e. the entire address number for each element).

```{r ex6}

```





