---
title: "AppEx 12: Hypothesis Testing"
subtitle: "Part 2"
date: "10/31/2022"
output: 
  html_document:
    theme: readable
---

## Resource for full examples in infer

https://infer.netlify.app/articles/observed_stat_examples.html



## Clone a repo + start a new project

Clone the **appex-12** repo on [GitHub](https://github.com/math118-fall2022) and start a new project in RStudio. 

We will live-code together, and then you must complete the remainder of this Application Exercise on your own!


## Data

We will continue working with the Asheville Airbnb data.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(infer)
```

```{r load-data,  message = F}
asheville <- read_csv("data/asheville.csv")
```

## LIVE CODE TOGETHER!

### Exercise 1

Suppose you are interested in whether at least half of the Airbnb listings in Asheville are more than $50 per guest per night. What would be your null and alternative hypotheses? There are two options! 

#### Option 1

#### Option 2


### Exercise 2

Simulate the null distribution to test your hypotheses. You can use 1000 reps for the in-class exercise.

#### Option 1 

```{r}
# don't forget to set a seed! 


```

#### Option 2

```{r}
#create variable to track price

```

```{r}
# don't forget to set a seed! 

```


### Exercise 3

What was your p-value? What decision do you make with respect to your 
hypotheses, and what conclusion do you make in the context of the research
problem?

#### Option 1

```{r}
#calculate observed statistic


# obtain p-value

```


#### Option 2

```{r}
#calculate observed statistic


# obtain p-value


```


## YOUR TURN!

### Exercise 4

Suppose you are interested in whether at most half of listings have a price per guest per night less than $50. How would your null and 
alternative hypotheses change in this case? Carry out the appropriate
hypothesis test (Option 1 and/or Option 2 above), and report your p-value, decision, and conclusion in context of the research problem.  **Use a seed of 1234 to conduct this hypothesis test.**


$H_0$: 

$H_a$: 


```{r ex4}
# your code here!

```

