---
title: "AppEx 19: Chipotle (cont.)"
date: "12/3/2022"
output: 
  html_document: 
    theme: readable
editor_options: 
  chunk_output_type: console
---

## Packages and Data

The packages you will need for this AppEx are listed below. You may need to install some of them. If so, install the appropriate packages in the Console!

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(sf) 
library(spData) 
library(leaflet)
```

Today's data set comes from the [Kaggle](https://www.kaggle.com/datasets/jeffreybraun/chipotle-locations?resource=download), and contains the locations of each Chipotle location in the US as of 07/27/2020. Someone scraped this data from Chipotle.

The variables are 

- `state`: county name
- `location`: number of registered voters
- `address`: number of individuals who have voted
- `latitude`: number of mail ballots returned and accepted
- `longitude`: number of mail ballots rejected

We will also use the `us_states` `sf` object from the `spData` library to map the contiguous US.

```{r}
chipotle <- read.csv("data/chipotle_stores.csv")
data("us_states")
```

## Exercises


### Exercise 1

Create an interactive visualization using `leaflet()` to display the Chipotle locations. Specifically, use the `add_markers()` function with appropriate arguments for `lng` and `lat`. You should use three functions in total!

```{r ex1}

```

### Exercise 2

Now let's try to create some summary data at the level of location. Group the `chipotle` data by both `state`  and `location`, then use the `summarise()` function to obtain the average `longitude` and `latitude` of each `location`, as well as the total number of stores at each `location`. Don't forget to use `ungroup()` at the end! 

Store this into a data frame called `chipotle_locations`. The first five observations should appear as follows:

```{r echo = F}
chipotle_locations <- chipotle %>% 
  group_by(state, location) %>%
  summarise(longitude = mean(longitude), latitude = mean(latitude),
            count = n()) %>%
  ungroup() 

chipotle_locations %>% 
  slice(1:5)
```

### Exercise 3

Now let's visualize the count of Chipotles at each `location` using `leaflet()`. Add circle markers using the function `addCircleMarkers()`, and make sure you have the following features for your markers:

  - The radius of the circles should be a function of the number Chipotles
  - The name of the `location` should appear when you click on the marker


### Exercise 4

Lastly, let's add a marker for the location of the Chipotle headquarters. The following code creates a custom marker in using an image of a burrito found online! I wanted to use the Chipotle logo, but I'm not sure if it's illegal to us it...

```{r icon}
burritoIcon <- makeIcon(
  iconUrl = "https://www.pngmart.com/files/5/Burrito-Transparent-PNG.png",
  iconWidth = 30, iconHeight = 30,
  iconAnchorX = 0, iconAnchorY = 0
)
```

Go online and find the longitude-latitude coordinates of the Chipotle headquarters. Add them into the following data frame:

```{r chipotle_hq, eval = F}
chipotle_hq <- data.frame(longitude = ____, latitude = ____)
```

Lastly, copy your code from Exercise 3 and paste it into the following code chunk. You will see another `addMarkers()` function that adds the `chipotle_hq` data to your leaflet visualization. Within this function, customize the marker as follows:

  - Set the longitude and latitude appropriately
  - The phrase "Chipotle HQ" should appear when you click on the marker
  - Set the optional `icon` argument of `addMarkers()` to the `burritoIcon`

```{r eval = F}
# paste code here %>%
  addMarkers(data = chipotle_hq, _________)
```
