---
title: "AE 08: Alinity COVID-19 Test"
name: "Your name"
date: "10/03/2022"
output: 
  html_document:
    theme: readable
---

## Clone a repo + start a new project

Go to the `appex-08-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. 

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

**Use R as a calculator for the exercises in this AE.** 

## Warm-up

Suppose in a class that 80% of the students identified as male and 20% identify as female. 60% of the male students are Econ majors and 80% of the female students are Econ majors. 

### Exercise 1

What is the probability that a randomly selected student in this class is an Econ major?

```{r ex-1}
```

### Exercise 2

Given that a student is an Econ major, what is the probability that they are male?

```{r ex-2}
```

## The Abbott Alinity Test

The global coronavirus pandemic illustrates the need for accurate testing of 
COVID-19, as its extreme infection rate poses a significant public health threat. 
Due to the time-sensitive nature of the situation, the FDA enacted emergency 
authorization of a number of serological tests for COVID-19. Full details of 
these tests may be found on its website 
[here](https://www.fda.gov/medical-devices/coronavirus-disease-2019-covid-19-emergency-use-authorizations-medical-devices/eua-authorized-serology-test-performance).

The Abbott Alinity test has an estimated sensitivity of 100% and specificity of 
99%. Let $D$ be the event that a patient has the disease, and let $T$ be the event that the test is positive for that disease.

- Prevalence: $P(D)$
- Sensitivity: $P(T | D)$
- Specificity: $P(T^c | D^c)$
- Positive predictive value: $P(D | T)$
- Negative predictive value: $P(D^c | T^c)$

### Exercise 3

Suppose the prevalence of COVID-19 in the general population is 2%. What are 
the positive and negative predictive values of the Abbott Alinity test?

```{r ex-3}
```

### Exercise 4

Now suppose the prevalence of COVID-19 in the general population is 8%. Repeat
the same calculations as in Exercise 3. Can you reuse code from Exercise 3
somehow?

```{r ex-4}
```

### Exercise 5

In the context of COVID-19, would you prefer to have a test with perfect 
sensitivity and good specificity or would you rather have a test with perfect 
specificity and good sensitivity? Explain your answer.
