---
title: "Spatial visualization: notes and live code"
output: html_document
date: "12/2/2022"
editor_options: 
  chunk_output_type: console
---

You may need to install some of these packages. If so, install the necessary packages in the Console!

```{r setup, message = F, warning = F}
library(tidyverse)
library(sf) # for spatial mapping
library(RColorBrewer) # for fun colors
```

## LECTURE NOTES

### Simple visualization

```{r fig.height = 2, eval = F}
nc <- st_read("data/nc_votes.shp", quiet = TRUE)

ggplot(___) + 
  ______ + 
  labs(title = "North Carolina counties")
```

### Choropleth map

```{r fig.height = 2, eval = F}
ggplot(nc) +
  geom_sf(aes(fill = ____)) + 
  labs(title = "Higher population counties have more votes cast",
       fill = "Total votes cast") 
```

### Customize colors

```{r fig.height = 2, eval = F}
ggplot(nc) +
  geom_sf(aes(fill = ____)) +
  scale_fill_gradient(low = "____", high = "____")
```


### More informative map?

```{r fig.height = 2, eval = F}
ggplot(nc) +
  geom_sf(aes(fill = ____)) +
  scale_fill_gradient(low = "____", high = "____") +
  labs(fill = "Votes cast per registered voter", 
       title = "Early vote turnout varies by county") 
#+ theme(legend.position = "____", legend.direction = "____")
```


## LIVE CODE

Recall `nycflights13` library which had all kinds information about all the flights out of NYC airports in 2013. In particular, the `flights` data frame contains the data for all flights that departed NYC, and `airports` contains useful metadata about the airports.

```{r message = F, warning = F}
library(spData) # for one of the datasets we will use
data("us_states")
library(nycflights13)
data(airports)
data(flights)
```

## Mapping states

```{r fig.height = 2}
## code here
```

## Mapping airports

```{r fig.height = 2}
## code here
```

## Wrangle

What if we wanted each point to represent the number of flights that flew to those airports from NYC? 

```{r}
## code here
```

## Mapping airport counts

```{r fig.height = 2}
## code here
```